A distance metric between histograms through the the Kolmogorov-Smirnov test statistic: specification, measures reference and example uses

### 2015-10-30 paper.tex
revisadas tabelas ateh os status

\section{Text}
\input{tables/textsGeneral}
\input{tables/textsDistances}
\input{tables/textsDistances2}
\input{tables/textsDistances2b}
\input{tables/textsDistances3}
\input{tables/textsDistances4}
\input{tables/textsDistances4b}
\section{Audio}
\input{tables/audioGeneral}
\input{tables/audioDistances}

\section{Music}
\input{tables/musicGeneral}
\input{tables/musicDistances}

\section{OS status}
\input{tables/osGeneral}
\input{tables/osDistances}

\section{Conclusions}\label{sec:conc}
The $c'$ metric is robust both to determine if
the distributions underlying the samples are the same
and to quantify the difference between probability distributions.
This work presents a set of benchmarks for making use $c'$.
Example analyses of real phenomena reveal patterns
without any of the training or clusterization usually involved in
classification routines.
Also, for making the analysis with other distributions
and data, or modifications on the measure of $c'$,
the rendering of this document is automatized in software.

>>>>

This section presents immediate results
drawn from the statistic $c$ when observed
in real samples.
The sample choices are arbitrary.

\section{Text}
This section exemplifies the use of $c$
in the detection of similarity between texts.
Each text $X$ was divided in two halves $X1$ and $X2$.
The set of known English words were considered as were 
the set of stopwords (words with reduced meaning such
as prepositions and articles).
Only the number of letters in each words was measured.
Three approaches were chosen: 1) the text was partitioned into 1000 pieces of equal number of characters, the mean of the word size of each piece is an element of the sample; 2) the text was partitioned into 1000 pieces of equal number of characters, the standard deviation of the word size is an element of the sample; 3) each word size is an element of the sample.
This last case yields a discrete probability distribution, which was approximated as a continuous variable and gave the greatest sensibility to text differences.
The overall result is the same: smaller differences between parts
of the same text.
Notice that the $c$ is often high within a same book.

\input{tables/textsGeneral}
\input{tables/textsDistances}
\input{tables/textsDistances2}
\input{tables/textsDistances2b}
\input{tables/textsDistances3}
\input{tables/textsDistances4}
\input{tables/textsDistances4b}

\FloatBarrier
\section{Audio}
This section presents $c$ values
drawn from audio for testing the sound system of the computer.
The PCM samples of the files were normalized to fit the interval
$[-1,1]$ to yield the samples labeled. The wavelet decomposition was performed with the Daubechies 8 Wavelet function.
The resulting values of the $c$ statistic reflect most of all the
different types of signals analysed:
PCM samples and wavelet decomposition coefficients in different leafs.
Among each type of signal, the type of sound is also reflected in the measures of $c$, with the noise having the highest values.

\input{tables/audioGeneral}
\input{tables/audioDistances}

\FloatBarrier
\section{Music}
This section presents measures of the $c$ statistic drawn
from the pitches of the notes of classical compositions.
The results reflect music history.
Take, for example, the first line, labeled ``Pale''.
The value of $c$ is increased with the exception of Beethoven
who, indeed, used modalism.
In the second line, labeled ``Bach1'',
the values of $c$ are ever increasing,
and previous $c$ (``Pale'' $\times$ ``Bach1'')
is only exceeded when ``Schoe`` is reached,
which reflects the non-tonal discourse of both
Palestrina and Sch\"onberg.

\input{tables/musicGeneral}
\input{tables/musicDistances}

\FloatBarrier
\section{OS status}
\input{tables/osGeneral}
\input{tables/osDistances}

\FloatBarrier
\section{Conclusions}\label{sec:conc}
The $c'$ metric is robust both to determine if
the distributions underlying the samples are the same
and to quantify the difference between probability distributions.
This work presents a set of benchmarks for making use $c'$.
Example analyses of real phenomena reveal patterns
without any of the training or clusterization usually involved in
classification routines.
Also, for making the analysis with other distributions
and data, or modifications on the measure of $c'$,
the rendering of this document is automatized in software.

### 2015-10-30 tables/musicGeneral.tex
terminada tabela com status do sistema operacional
revisadas tabelas ateh os status

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l | c | c |}\hline
label & description & events \\\hline
Pale & Sanctus 69 from G. P. da Palestrina  & 719.00 \\\hline
Bach1 & BWV735 from J. S. Bach  & 236.00 \\\hline
Bach2 & BWV648 from J. S. Bach  & 272.00 \\\hline
Moza1 & K80 from W. A. Mozart  & 538.00 \\\hline
Moza2 & K458 from W. A. Mozart  & 4218.00 \\\hline
Beet1 & Opus 18, n1, mov. 3 from L. van Beethoven  & 1289.00 \\\hline
Beet2 & Opus 132 from L. van Beethoven  & 17884.00 \\\hline
Shoe & Opus 19, mov. 2 from A. Sh\"nberg  & 102.00 \\\hline
\end{tabular}
\caption{General description of the music data used for the $c'$ values of the next table. Each event is a midi value of a note pitch.}
\end{center}
\end{table}

>>>>

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l | c | c |}\hline
label & description & events \\\hline
Pale & Sanctus 69 from G. P. da Palestrina  & 719.00 \\\hline
Bach1 & BWV735 from J. S. Bach  & 236.00 \\\hline
Bach2 & BWV648 from J. S. Bach  & 272.00 \\\hline
Moza1 & K80 from W. A. Mozart  & 538.00 \\\hline
Moza2 & K458 from W. A. Mozart  & 4218.00 \\\hline
Beet1 & Opus 18, n1, mov. 3 from L. van Beethoven  & 1289.00 \\\hline
Beet2 & Opus 132 from L. van Beethoven  & 17884.00 \\\hline
Sch\"on & Opus 19, mov. 2 from A. Sch\"onberg  & 102.00 \\\hline
\end{tabular}
\caption{General description of the music data used for the $c$ values of the next table. Each event is a midi value of a note pitch. Samples where chosen to reflect music history timeline. Works by the same composer were chosen among the first and last 10\% of all he produced.}
\end{center}
\end{table}

### 2015-10-30 tables/tabNormNull.tex
revisadas tabelas ateh os status

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c |}\hline
$\alpha N_c$ & $\alpha$ & $c(\alpha)$ & $|C_1(\alpha)|$ & $|C_2(\alpha)|$ & $|C_3(\alpha)|$ \\\hline
10.0 & 0.100 & 1.22 & 3 & 2 & 1 \\\hline
5.0 & 0.050 & 1.36 & 0 & 1 & 0 \\\hline
2.5 & 0.025 & 1.48 & 0 & 0 & 0 \\\hline
1.0 & 0.010 & 1.63 & 0 & 0 & 0 \\\hline
0.5 & 0.005 & 1.73 & 0 & 0 & 0 \\\hline
0.1 & 0.001 & 1.95 & 0 & 0 & 0 \\\hline
\end{tabular}
\caption{The theoretical maximum number $\alpha N_c$ of rejections
of the null hypothesis for significance levels $\alpha$.
The $c_1$ values were calculated using simulations of normal distributions with $\mu=0$ and $\sigma=1$.
The $c_2$ values were calculated using simulations of normal distributions with $\mu=3$ and $\sigma=2$.
The $c_3$ values were calculated using simulations of normal distributions with $\mu=6$ and $\sigma=3$.
Over all $N_c$ comparisons,
 $\mu(c_1)=0.6979$ and $\sigma(c_1)=0.2339$,
 $\mu(c_2)=0.6838$ and $\sigma(c_2)=0.2697$,
 $\mu(c_3)=0.6809$ and $\sigma(c_3)=0.2220$ .
}
\end{center}
\end{table}

>>>>

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c |}\hline
$\alpha N_c$ & $\alpha$ & $c(\alpha)$ & $|C_1(\alpha)|$ & $|C_2(\alpha)|$ & $|C_3(\alpha)|$ \\\hline
10.0 & 0.100 & 1.22 & 3 & 3 & 9 \\\hline
5.0 & 0.050 & 1.36 & 1 & 2 & 3 \\\hline
2.5 & 0.025 & 1.48 & 0 & 0 & 2 \\\hline
1.0 & 0.010 & 1.63 & 0 & 0 & 0 \\\hline
0.5 & 0.005 & 1.73 & 0 & 0 & 0 \\\hline
0.1 & 0.001 & 1.95 & 0 & 0 & 0 \\\hline
\end{tabular}
\caption{The theoretical maximum number $\alpha N_c$ of rejections
of the null hypothesis for significance levels $\alpha$.
The $c_1$ values were calculated using simulations of normal distributions with $\mu=0$ and $\sigma=1$.
The $c_2$ values were calculated using simulations of normal distributions with $\mu=3$ and $\sigma=2$.
The $c_3$ values were calculated using simulations of normal distributions with $\mu=6$ and $\sigma=3$.
Over all $N_c$ comparisons,
 $\mu(c_1)=0.6527$ and $\sigma(c_1)=0.2398$,
 $\mu(c_2)=0.6889$ and $\sigma(c_2)=0.2397$,
 $\mu(c_3)=0.7184$ and $\sigma(c_3)=0.2811$ .
}
\end{center}
\end{table}

### 2015-10-30 tables/tabPowerNull.tex
revisadas tabelas ateh os status

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c | c | c |}\hline
$\alpha N_c$ & $\alpha$ & $c(\alpha)$ & $|C_1(\alpha)|$ & $|C_2(\alpha)|$ & $|C_3(\alpha)|$ & $|C_4(\alpha)|$ & $|C_5(\alpha)|$ \\\hline
10.0 & 0.100 & 1.22 & 5 & 8 & 6 & 8 & 3 \\\hline
5.0 & 0.050 & 1.36 & 3 & 4 & 1 & 1 & 3 \\\hline
2.5 & 0.025 & 1.48 & 2 & 1 & 0 & 1 & 2 \\\hline
1.0 & 0.010 & 1.63 & 0 & 0 & 0 & 0 & 1 \\\hline
0.5 & 0.005 & 1.73 & 0 & 0 & 0 & 0 & 0 \\\hline
0.1 & 0.001 & 1.95 & 0 & 0 & 0 & 0 & 0 \\\hline
\end{tabular}
\caption{The theoretical maximum number $\alpha N_c$ of rejections
of the null hypothesis for critical values of $\alpha$.
The $c_1$ values were calculated using simulations of power functions distributions with $a=0.3$.
The $c_2$ values were calculated using simulations of power functions distributions with $a=1$.
The $c_3$ values were calculated using simulations of power functions distributions with $a=2$.
The $c_4$ values were calculated using simulations of power functions distributions with $a=3$.
The $c_5$ values were calculated using simulations of power functions distributions with $a=4$.
Over all $N_c$ comparisons,
 $\mu(c_1)=0.7160$ and $\sigma(c_1)=0.2645$,
 $\mu(c_2)=0.8052$ and $\sigma(c_2)=0.2389$,
 $\mu(c_3)=0.7424$ and $\sigma(c_3)=0.2458$ .
 $\mu(c_4)=0.7580$ and $\sigma(c_4)=0.2538$ .
 $\mu(c_5)=0.7265$ and $\sigma(c_5)=0.2655$ .
}
\end{center}
\end{table}

>>>>

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c | c | c |}\hline
$\alpha N_c$ & $\alpha$ & $c(\alpha)$ & $|C_1(\alpha)|$ & $|C_2(\alpha)|$ & $|C_3(\alpha)|$ & $|C_4(\alpha)|$ & $|C_5(\alpha)|$ \\\hline
10.0 & 0.100 & 1.22 & 5 & 5 & 7 & 5 & 5 \\\hline
5.0 & 0.050 & 1.36 & 2 & 3 & 5 & 1 & 1 \\\hline
2.5 & 0.025 & 1.48 & 1 & 2 & 1 & 1 & 0 \\\hline
1.0 & 0.010 & 1.63 & 0 & 0 & 0 & 1 & 0 \\\hline
0.5 & 0.005 & 1.73 & 0 & 0 & 0 & 0 & 0 \\\hline
0.1 & 0.001 & 1.95 & 0 & 0 & 0 & 0 & 0 \\\hline
\end{tabular}
\caption{The theoretical maximum number $\alpha N_c$ of rejections
of the null hypothesis for critical values of $\alpha$.
The $c_1$ values were calculated using simulations of power functions distributions with $a=0.3$.
The $c_2$ values were calculated using simulations of power functions distributions with $a=1$.
The $c_3$ values were calculated using simulations of power functions distributions with $a=2$.
The $c_4$ values were calculated using simulations of power functions distributions with $a=3$.
The $c_5$ values were calculated using simulations of power functions distributions with $a=4$.
Over all $N_c$ comparisons,
 $\mu(c_1)=0.6883$ and $\sigma(c_1)=0.2617$,
 $\mu(c_2)=0.7701$ and $\sigma(c_2)=0.2432$,
 $\mu(c_3)=0.7871$ and $\sigma(c_3)=0.2592$ .
 $\mu(c_4)=0.7721$ and $\sigma(c_4)=0.2587$ .
 $\mu(c_5)=0.7008$ and $\sigma(c_5)=0.2463$ .
}
\end{center}
\end{table}

### 2015-10-30 tables/tabUniformNull.tex
revisadas tabelas ateh os status

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c |}\hline
$\alpha N_c$ & $\alpha$ & $c(\alpha)$ & $|C_1(\alpha)|$ & $|C_2(\alpha)|$ & $|C_3(\alpha)|$ \\\hline
10.0 & 0.100 & 1.22 & 3 & 5 & 11 \\\hline
5.0 & 0.050 & 1.36 & 2 & 2 & 9 \\\hline
2.5 & 0.025 & 1.48 & 0 & 2 & 3 \\\hline
1.0 & 0.010 & 1.63 & 0 & 1 & 2 \\\hline
0.5 & 0.005 & 1.73 & 0 & 1 & 1 \\\hline
0.1 & 0.001 & 1.95 & 0 & 0 & 0 \\\hline
\end{tabular}
\caption{The theoretical maximum number $\alpha N_c$ of rejections
of the null hypothesis for critical values of $\alpha$.
The $c_1$ values were calculated using simulations of uniform distributions within $[0,1)$.
The $c_2$ values were calculated using simulations of uniform distributions within $[2,6)$.
The $c_3$ values were calculated using simulations of uniform distributions with $\mu=4$ and $\sigma=10$.
Over all $N_c$ comparisons,
 $\mu(c_1)=0.7547$ and $\sigma(c_1)=0.2460$,
 $\mu(c_2)=0.7464$ and $\sigma(c_2)=0.2803$,
 $\mu(c_3)=0.8027$ and $\sigma(c_3)=0.3119$ .
}
\end{center}
\end{table}

>>>>

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c |}\hline
$\alpha N_c$ & $\alpha$ & $c(\alpha)$ & $|C_1(\alpha)|$ & $|C_2(\alpha)|$ & $|C_3(\alpha)|$ \\\hline
10.0 & 0.100 & 1.22 & 5 & 3 & 3 \\\hline
5.0 & 0.050 & 1.36 & 4 & 1 & 2 \\\hline
2.5 & 0.025 & 1.48 & 2 & 1 & 1 \\\hline
1.0 & 0.010 & 1.63 & 1 & 1 & 0 \\\hline
0.5 & 0.005 & 1.73 & 0 & 0 & 0 \\\hline
0.1 & 0.001 & 1.95 & 0 & 0 & 0 \\\hline
\end{tabular}
\caption{The theoretical maximum number $\alpha N_c$ of rejections
of the null hypothesis for critical values of $\alpha$.
The $c_1$ values were calculated using simulations of uniform distributions within $[0,1)$.
The $c_2$ values were calculated using simulations of uniform distributions within $[2,6)$.
The $c_3$ values were calculated using simulations of uniform distributions with $\mu=4$ and $\sigma=10$.
Over all $N_c$ comparisons,
 $\mu(c_1)=0.7656$ and $\sigma(c_1)=0.2552$,
 $\mu(c_2)=0.7520$ and $\sigma(c_2)=0.2421$,
 $\mu(c_3)=0.7077$ and $\sigma(c_3)=0.2250$ .
}
\end{center}
\end{table}

### 2015-10-30 tables/tabWeibullNull.tex
revisadas tabelas ateh os status

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c | c |}\hline
$\alpha N_c$ & $\alpha$ & $c(\alpha)$ & $|C_1(\alpha)|$ & $|C_2(\alpha)|$ & $|C_3(\alpha)|$ & $|C_4(\alpha)|$ \\\hline
10.0 & 0.100 & 1.22 & 0 & 4 & 8 & 6 \\\hline
5.0 & 0.050 & 1.36 & 0 & 1 & 2 & 1 \\\hline
2.5 & 0.025 & 1.48 & 0 & 0 & 1 & 1 \\\hline
1.0 & 0.010 & 1.63 & 0 & 0 & 1 & 0 \\\hline
0.5 & 0.005 & 1.73 & 0 & 0 & 1 & 0 \\\hline
0.1 & 0.001 & 1.95 & 0 & 0 & 0 & 0 \\\hline
\end{tabular}
\caption{The theoretical maximum number $\alpha N_c$ of rejections
of the null hypothesis for critical values of $\alpha$.
The $c_1$ values were calculated using simulations of 1-parameter Weibull distributions with $a=0.1$.
The $c_2$ values were calculated using simulations of 1-parameter Weibull distributions with $a=2$.
The $c_3$ values were calculated using simulations of 1-parameter Weibull distributions with $a=4$.
Over all $N_c$ comparisons,
The $N_o$ values of $c_4$ were calculated using simulations of
 1-parameter Weibull distributions with $a=6$.
Over all $N_c$ comparisons,
 $\mu(c_1)=0.0577$ and $\sigma(c_1)=0.0357$,
 $\mu(c_2)=0.7518$ and $\sigma(c_2)=0.2305$,
 $\mu(c_3)=0.7281$ and $\sigma(c_3)=0.2958$ .
 $\mu(c_4)=0.7189$ and $\sigma(c_4)=0.2625$ .
}
\end{center}
\end{table}

>>>>

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c | c |}\hline
$\alpha N_c$ & $\alpha$ & $c(\alpha)$ & $|C_1(\alpha)|$ & $|C_2(\alpha)|$ & $|C_3(\alpha)|$ & $|C_4(\alpha)|$ \\\hline
10.0 & 0.100 & 1.22 & 0 & 2 & 0 & 4 \\\hline
5.0 & 0.050 & 1.36 & 0 & 1 & 0 & 1 \\\hline
2.5 & 0.025 & 1.48 & 0 & 0 & 0 & 0 \\\hline
1.0 & 0.010 & 1.63 & 0 & 0 & 0 & 0 \\\hline
0.5 & 0.005 & 1.73 & 0 & 0 & 0 & 0 \\\hline
0.1 & 0.001 & 1.95 & 0 & 0 & 0 & 0 \\\hline
\end{tabular}
\caption{The theoretical maximum number $\alpha N_c$ of rejections
of the null hypothesis for critical values of $\alpha$.
The $c_1$ values were calculated using simulations of 1-parameter Weibull distributions with $a=0.1$.
The $c_2$ values were calculated using simulations of 1-parameter Weibull distributions with $a=2$.
The $c_3$ values were calculated using simulations of 1-parameter Weibull distributions with $a=4$.
Over all $N_c$ comparisons,
The $N_o$ values of $c_4$ were calculated using simulations of
 1-parameter Weibull distributions with $a=6$.
Over all $N_c$ comparisons,
 $\mu(c_1)=0.0548$ and $\sigma(c_1)=0.0314$,
 $\mu(c_2)=0.6878$ and $\sigma(c_2)=0.2352$,
 $\mu(c_3)=0.7030$ and $\sigma(c_3)=0.2037$ .
 $\mu(c_4)=0.7350$ and $\sigma(c_4)=0.2510$ .
}
\end{center}
\end{table}

### 2015-10-30 tables/textsGeneral.tex
revisadas tabelas ateh os status

\begin{table*}[h!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c | c | c | c |}\hline
label & description & chars & tokens & sents & $\mu(kw)$ & $\sigma(kw)$ & $\mu(sw)$ & $\sigma(sw)$ \\\hline
H,H1,H2 & Hamlet by Shakespeare & 162881 & 37360 & 3106 & 3.549 & 1.762 & 2.721 & 1.011 \\\hline
B,B1,B2 & King James Version of the Holly Bible & 4332554 & 1010654 & 30103 & 3.745 & 1.711 & 2.927 & 1.044 \\\hline
M,M1,M2 & Moby Dick by Herman Melville & 1242990 & 260819 & 10059 & 4.105 & 2.184 & 2.847 & 1.096 \\\hline
E,E1,E2 & Esa\'u e Jac\'o from Machado de Assis & 355706 & 88472 & 3822 & 2.186 & 1.376 & 1.486 & 0.502 \\\hline
\end{tabular}
\caption{General description of the texts used to exemplify the use of the $c'$ values. Individual values of number of characters, tokens, sentences give context. Mean and standard deviation of the size of known words $kw$ and of the stopwords $st$ used are used in next table for comparison through $c'$. This table holds has the only purpose of contextualizing next table. Numbers in the labels indicate first and second half of the corresponding text in the next tables.}
\end{center}
\end{table*}

>>>>

\begin{table*}[h!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c | c | c | c |}\hline
label & description & chars & tokens & sentences & $\mu(kw)$ & $\sigma(kw)$ & $\mu(sw)$ & $\sigma(sw)$ \\\hline
H,H1,H2 & Hamlet by Shakespeare & 162881 & 37360 & 3106 & 3.549 & 1.762 & 2.721 & 1.011 \\\hline
B,B1,B2 & King James Version of the Holly Bible & 4332554 & 1010654 & 30103 & 3.745 & 1.711 & 2.927 & 1.044 \\\hline
M,M1,M2 & Moby Dick by Herman Melville & 1242990 & 260819 & 10059 & 4.105 & 2.184 & 2.847 & 1.096 \\\hline
E,E1,E2 & Esa\'u e Jac\'o from Machado de Assis & 355706 & 88472 & 3822 & 2.186 & 1.376 & 1.486 & 0.502 \\\hline
\end{tabular}
\caption{General description of the texts used to exemplify the use of the $c$ statistic.
Individual values of number of characters, tokens, sentences give context.
Mean and standard deviation of the size of known words $kw$ and of the stopwords
$st$ are used in next tables.
Numbers in the labels indicate first and second half of the corresponding text in the next tables.}
\end{center}
\end{table*}
